---
title: "Abalone Intro Exercise"
author: "Evan Hope"
date: "2024-03-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this exercise is to display my introductory knowledge of reading
in a data set, appending a variable to the set, setting up a workflow, creating 
a recipe, splitting data into training and test sets, and using the recipe and training set to help create predictions of the age of an abalone.

Below is a walk through of my work.


Before I begin, I will download the packages needed to conduct data manipulation
and modeling. 

```{r}
install.packages("tidyverse")
```

```{r}
install.packages("tidymodels")
```

```{r}
install.packages("ISLR")
```

```{r}
library(tidyverse)
library(tidymodels)
library(ISLR)
library(dplyr)
```

Now that the necessary packages are ready to go, I will now read in the abalone 
data set.

```{r}
abalone_data = read.csv("C:/Users/Ordai/OneDrive/Desktop/School/Personal Projects/Abalone Intro Exercise/abalone.csv")
```

I will slightly modify the data set by adding an "age" variable. 

```{r}
abalone_data2 <- mutate(abalone_data, age = rings + 1.5)
# The age (in years) of an abalone can be quickly estimated by adding 1.5 to
# the number of rings found within the abalone. Hence why age = rings + 1.5

glimpse(abalone_data2)
```

Note that my end goal of this exercise is to predict the age of an abalone by 
creating a recipe that uses a combination of variables EXCLUDING the "rings" 
variable. The "rings" variable as well as the formula used above (rings+1.5=age)
will only be used as a reference to see how close my recipe comes to predicting 
the actual age (rings + 1.5).






